Copulas and their potential for ecology

### 2019-12-15 AER_Responses.tex
some more working on responses for AER
a bit more progress on responses to AER refs

>>>We were not entirely sure what the referee was getting at here, but we
thank the referee for the comment because it spurred us to create a new table, 
table X, which summarizes the properties of the different families.
%***DAN: Replace X with whatever the number of the new table turns out to be-->

@@

@@

@@

>>>>

>>>We were not entirely sure what the referee was getting at here, but we
thank the referee for the comment because it spurred us to create a new table, 
table 2, which summarizes the properties of the different families.

@@

>>>Thank you for asking this question, as it prompted us to we add a new paragraph
on this topic to the Discussion. The new paragraph begins "Data requirements for 
copula methods..." and is the second-to-last paragraph of the main text.

@@

>>>Thanks for the suggestion, we have made this change.

@@

>>>Thank you for this suggestion. We have considered at various times the possibility 
of splitting the manuscript up
and/or removing some of its content. In the end we decided it is best to keep it all together,
for several reasons. First, the parts of the manuscript (theoretical and empirical, the three
research question, the multiple case studies) reinforce each other, providing better evidence
for the points we are making than any of these parts would on its own. Second,
we selected datasets in an attempt to span multiple disciplines of ecology.
Removing some now would create a less representative collection of datasets across the field, and would weaken
our claim that copula structure seems to be a common phenomenon, broadly, across ecology. 
Finally, we selected datasets prior to analysis and we present all results. This was done
to be sure to fairly evaluate the hypothesis that non-Gaussian copula structures are
common in ecology. We did *not* carry out additional analyses on other datasets and
only present the results with interesting copula structure. We would prefer not to 
modify the manuscript in such a way as to invalidate this strength of our work.

### 2019-12-15 Paper.Rmd
a bit more progress on responses to AER refs

```{r tab_summary_copula_info, echo=F, results="asis",message=F}
library(kableExtra)
library(dplyr)
tab_copula_info<-as.data.frame(matrix(NA,nrow=16,ncol=4))
colnames(tab_copula_info)<-c("Copula family","No. of parameters","Stronger dependent tail", "Reference")

@@

tab_copula_info$`Stronger dependent tail`<-c("None",
                                             "Lower","Upper",
                                             "Upper","Lower",
                                             "Upper","Lower",
                                             "None",
                                             "Any (depending on the parameters)","Any (depending on the parameters)",
                                             "Upper","Lower",
                                             "Any (depending on the parameters)","Any (depending on the parameters)",
                                             "Either upper or none (depending on the parameters)","Either lower or none (depending on the parameters)")

@@

knitr::kable(tab_copula_info, 
             format="latex", align="l",escape=F,
             caption = "Summary table for the sixteen copula families we used (see section \\ref{Background}). \\label{tab_copula_info}", 
             booktabs = T,linesep = "\\addlinespace") %>%
             column_spec(2,width="2 cm")
```

@@

We addressed Q1 first via a model selection procedure in which several
families of copulas were fitted to 
our ecological datasets and fits were compared via the Akaike and Bayesian 
Information Criteria (AIC and BIC). One of the fitted copulas was the normal
copula, making possible comparisons of the degree to which the normal copula versus
other copula families were good descriptions of data.

@@

Having demonstrated that non-normal copula structure and asymmetric tail associations are 
addressed Q2 by exploring, using models, three possible mechanisms
by which these phenomena may arise. 
Models presented are initial explorations, only, of whether the proposed mechanisms have the 
potential to explain observed patterns. As such, simple models were used. 
Comprehensive explorations of
model parameter space and alternative model formulations were left for future work.

@@

We addressed Q3 by exploring, using both data and models, three hypothesized
ecological consequences of non-normal copula structure and asymmetric tail
associations. As for Q2, the models we used are intended as initial explorations, only,
of whether copula structure may have a meaningful influence on important ecological
phenomena. Therefore only simple models are used, and comprehensive explorations of the 
sensitivity of results to model strucuture and parameters are left for future work.

>>>>

```{r tab_summary_copula_info, echo=F, results="asis",message=F}
library(kableExtra)
library(dplyr)
tab_copula_info<-as.data.frame(matrix(NA,nrow=16,ncol=4))
colnames(tab_copula_info)<-c("Copula family","No. of parameters","Tails with stronger association", "Reference")

@@

tab_copula_info$`Tails with stronger association`<-c("Neither (symmetric)",
                                             "Lower","Upper",
                                             "Upper","Lower",
                                             "Upper","Lower",
                                             "Neither (symmetric)",
                                             "Either (depending on params)","Either (depending on params)",
                                             "Upper","Lower",
                                             "Either (depending on params)","Either (depending on params)",
                                             "Upper or symmetric (depending on paras)","Lower or symmetric (depending on params)")

@@

knitr::kable(tab_copula_info, 
             format="latex", align="l",escape=F,
             caption = "Summary table for the sixteen copula families we used (see section \\ref{Background}). For the parameters we consider, all copula families model positive association between variables. \\label{tab_copula_info}", 
             booktabs = T,linesep = "\\addlinespace") %>%
             column_spec(2,width="2 cm")
```

@@

Recall that Q1 has three parts: Do datasets
in ecology have dependence structure distinct from that of a multivariate 
Gaussian/normal
distribution (here called non-normal copula structure)? Do positvely associated 
ecological variables show tail 
associations distinct from those of a normal distribution? And in particular
do they show asymmetric tail associations? 
We addressed Q1 first via a model selection procedure in which several
families of copulas were fitted to 
our ecological datasets and fits were compared via the Akaike and Bayesian 
Information Criteria (AIC and BIC). One of the fitted copulas was the normal
copula, making possible comparisons of the degree to which the normal copula versus
other copula families were good descriptions of data.

@@

Recall that Q2 is: what are some possible causes/mechanisms of non-normal
copula structure and asymmetric tail associations in ecology?
Having demonstrated that non-normal copula structure and asymmetric tail associations are 
common in ecology, we addressed Q2 by exploring, using models, three possible mechanisms
by which these phenomena may arise. 
Models presented are initial explorations, only, of whether the proposed mechanisms have the 
potential to explain observed patterns. As such, simple models were used. 
Comprehensive explorations of
model parameter space and alternative model formulations were left for future work.

@@

Recall that Q3 is: what are the consequences of non-normal copula
structure and asymmetric tail associations for ecological understanding and
applications?
We addressed Q3 by exploring, using both data and models, three hypothesized
ecological consequences of non-normal copula structure and asymmetric tail
associations. As for Q2, the models we used are intended as initial explorations, only,
of whether copula structure may have a meaningful influence on important ecological
phenomena. Therefore only simple models are used, and comprehensive explorations of the 
sensitivity of results to model strucuture and parameters are left for future work.
